A Bidirected String Graph Model for Genome Assembly

### 2012-12-05 presentation/paper.tex
Introduction

\title{A Bidirected String Graph Model for Genome Assembly}
\author{Eric Biggers\footnote{Macalester College}}

@@

%\begin{abstract}
%Sequencing an organism's genome is an important step to further genetic research
%on that organism.  However, current sequencing machines can only sequence very
%short fragments of DNA that originate from random positions on the original
%genome, assembly software such as Allpaths-LG can be used to assemble these
%fragments into longer contiguous sequences.  
%\end{abstract}

@@

\subsection{Definition of a genome}
\subsection{Overall workflow for genome sequencing}
\subsection{Genome assemblers}

>>>>

\title{A Bidirected String Graph Model for Genome Assembly}
\author{Eric Biggers}

@@

%\begin{abstract}
%Sequencing an organism's genome is an important step to further genetic research
%on that organism.  However, current sequencing machines can only sequence very
%short fragments of DNA that originate from random positions on the original
%genome, assembly software such as Allpaths-LG can be used to assemble these
%fragments into longer contiguous sequences.  
%\end{abstract}
\newcommand\Base[1]{{\tt #1}}

@@

\subsection{Definition of a genome}

Every living organism contains a collection of hereditary material called a {\em
genome}.  Typically, each individual cell of a multicellular organism contains a
complete copy of the genome.  A genome consists of one or more DNA ({\it
deoxyribonucleic acid}) molecules in the form of {\it chromosomes}.  DNA is a
double-stranded molecule, each strand of which is a polymer of {\em
nucleotides}.  A nucleotide, together with the corresponding nucleotide on the
opposite strand, is called a {\it base pair}.  In DNA, there are four possible
nucleotides: adenine, cytosine, guanine, and thymine, which are abbreviated as
\Base{A}, \Base{C}, \Base{G}, and \Base{T}, respectively.  \Base{A} always pairs
with T on the other strand, and \Base{C} always pairs with \Base{G} on the other
strand.  In addition, each strand has a direction, and the two strands of a DNA
molecule are directed opposite one another.  Furthermore, a DNA molecule may be
circular (as is usually the case for bacterial chromosomes), in which case the
molecule has no ends.

Therefore, for the purposes of this paper, a genome is defined as a set of
possibly circular, dual-stranded strings of the symbols \Base{A}, \Base{C},
\Base{G}, and \Base{T}, where dual-stranded has the meaning indicated above (the
two strands run in opposite directions, and the pairings \Base{A}---\Base{T} and
\Base{C}---\Base{G} always hold.)

This definition does not account for the fact that bacterial chromosomes are
circular, nor does it account for the fact that diploid and polyploid organisms
contain multiple, slightly-differing copies of each chromosome.  For the latter
case, the above definition will consider each copy of the chromosome to be a
completely different string in the genome.

Genomes vary widely in size, from several thousand base pairs to many billion
base pairs, depending on the organism.  (See Table \ref{tab:GenomeSizes}).

\begin{table}[H]
	\begin{center}
		\begin{tabular}{|l|p{4.5cm}|}
			%\hline
			%$\phi$X174 bacteriophage &  $\sim$5,000 bp \\
			%\hline
			%Human mitochondrion &  $\sim$16,000 bp \\
			\hline
			{\bf Organism} & {\bf Approximate genome size (in base
			pairs)} \\
			\hline
			{\it Variola} (smallpox) virus &  186,000 \\
			\hline
			{\it E. coli} (a common bacterium) & 4,600,000 \\
			\hline
			{\it Ananas comosus} (domesticated pineapple) & 500,000,000 \\
			\hline
			{\it Zea mays} (domesticated corn) & 2,000,000,000 \\
			\hline
			{\it Homo sapiens} (human)       &  3,200,000,000 \\
			\hline
		\end{tabular}
	\end{center}
	\caption{Approximate genome sizes of selected organisms}
	\label{tab:GenomeSizes}
\end{table}

\subsection{Genome sequencing}

{\it DNA sequencing} is the process of determining the base pair sequence of one
or more DNA molecules.  {\it Genome sequencing} is the use of DNA sequencing to
determine the base pair sequence of an entire genome.

\subsection{Overall workflow for genome sequencing}
\subsection{Genome assemblers}
